Quantifying the ease of viral escape from broad and narrow antibodies to influenza hemagglutinin

### 2017-10-15 paper/manuscript.tex
included all figures in LaTex

\subsection*{Quantifying the antigenic effects of all mutations selected by each antibody}
We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and analyzed the data to calculate the fraction of virions carrying each mutation that survived neutralization. 
All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all amino-acid mutants to the HA of the A/WSN/1933 (H1N1) strain~\citep{doud2016accurate}.
We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and we also re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
The correlations among replicates are in Figure~\ref{suppfig:corr}.
For the remainder of this paper, we will refer to the median antigenic effect of each mutation across replicates.

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figs/avgfracsurvive.pdf}}
\caption{
\label{fig:avgfracsurvive}
{\bf The strain-specific and broad anti-receptor-binding-pocket antibodies select mutations with large antigenic effects, but the broad anti-stalk antibodies only select mutations with small effects.}
The height of the line at each site indicates the excess fraction of virions with a mutation at that site that survive the antibody selection, averaging across all amino-acid mutations at the site (see Equation~\ref{eq:avgfracsurvive}).
It is immediately obvious that there are multiple large-effect mutations for H17L19, H17L10, H17L7, and S139/1 -- but not any large effect mutations for FI6v3 and C179.
Sites are labeled in H3 numbering.
Figure~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the single most strongly selected amino-acid mutation at each site.
Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo} show data for all mutations for each antibody using logo plots.
}
\end{figure}

It is immediately obvious that the narrow strain-specific antibodies and the antibody targeting HA's receptor-binding pocket (S139/1) select mutations with large antigenic effects.
Figure~\ref{fig:avgfracsurvive} shows that for all four of these antibodies, there are multiple sites in HA where mutations enable a substantial fraction of virions to survive high antibody concentrations.
Specifically, there are mutations that enable over a third of the virions to evade antibody at concentrations where virtually all the wildtype virions are neutralized (Figure~\ref{suppfig:maxfracsurvive}).
Therefore, the virus can escape these four antibodies with the type of large-effect single amino-acid mutations that are thought to underly to influenza antigenic drift.

@@

>>>>

\subsection*{Quantifying the antigenic effects of all mutations selected by each antibody}
We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and calculated the fraction of virions carrying each mutation that survived neutralization. 
All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all single amino-acid mutations to HA~\citep{doud2016accurate}.
We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies, and we re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
The correlations among replicates are in Figure~\ref{suppfig:corr}.
For the remainder of this paper, we will refer to the median antigenic effect of each mutation across replicates.

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figs/avgfracsurvive.pdf}}
\caption{
\label{fig:avgfracsurvive}
{\bf The strain-specific and broad anti-receptor-binding-site antibodies select mutations with large antigenic effects, but the broad anti-stalk antibodies only select small-effect mutations.}
The height of the lines indicate the excess fraction of virions with a mutation at that site that survive the antibody, averaging across all amino-acid mutations at each site (see Equation~\ref{eq:avgfracsurvive}).
There are multiple sites of large-effect mutations for H17L19, H17L10, H17L7, and S139/1 -- but not any for FI6v3 and C179.
Figure~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the most strongly selected amino-acid mutation at each site.
Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo} show data for all mutations for each antibody using logo plots.
Sites are labeled in H3 numbering.
}
\end{figure}

It is immediately obvious that the narrow strain-specific antibodies and the antibody targeting HA's receptor-binding pocket (S139/1) select mutations with large antigenic effects.
For all four of these antibodies, there are multiple sites in HA where mutations enable a substantial fraction of virions to survive high antibody concentrations (Figure~\ref{fig:avgfracsurvive}).
There are mutations that enable over a third of virions to evade these antibodies at concentrations where virtually all wildtype virions are neutralized (Figure~\ref{suppfig:maxfracsurvive}).
Therefore, the virus can escape these four antibodies with the sort of large-effect single amino-acid mutations that characterize traditional influenza antigenic drift~\citep{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.

@@

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figs/stalkAbs_zoomed/stalkAbs_zoomed.pdf}}
\caption{
\label{fig:stalkAbs_zoomed}
CAPTION
}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=\textwidth]{figs/FI6v3mutant_neutcurves/FI6v3_mutant_neutcurves.pdf}}
\caption{
\label{fig:FI6v3neutcurves}
CAPTION
}
\end{figure}
